---
title: "R Notebook"
output: html_notebook
---
#Plot the 30-day mortality rates for heart attack
Goal is to use only clear R without any additional packages
##Read & check data
```{r}
outcome <- read.csv("outcome-of-care-measures.csv", colClasses = "character")
head(outcome)

```
##Make a histogramm

```{r}
outcome[, 11] <- as.numeric(outcome[, 11])
hist(outcome[, 11])
```

# Finding the best hospital in a state
##Function
```{r}
best <- function(state, outcome) {
## Read outcome data
  data <- read.csv("outcome-of-care-measures.csv", colClasses = "character")

## Check that state and outcome are valid
  if(!state %in% data$State) stop("invalid state")
  outcomes = c("heart attack","heart failure","pneumonia")
  if(!outcome %in% outcomes) stop("invalid outcome")
  data <- data[data$State == state,]
  if(match(outcome,outcomes)==1) {
    col_num <- 11
  }
  else if(match(outcome,outcomes)==2) {
    col_num <- 17
  }
  else col_num <- 23
  data[data[,col_num]=="Not Available",] <- NA
  data <- data[!is.na(data[,col_num]),]
  data[,col_num] <- as.numeric(data[,col_num])

  result <- data[data[,col_num] == min(data[,col_num], na.rm=T),]
  result <- result[order(result[,col_num]),]
  

## Return hospital name in that state with lowest 30-day death
## rate
  result[1,2]
}

```
##Test function
```{r}
best("TX", "heart attack")
best("TX", "heart failure")
best("MD", "heart attack")
best("MD", "pneumonia")
```
##Test incorrect input
```{r}
best("BB", "heart attack")
```
```{r}
best("NY", "hert attack")
```


#Ranking hospitals by outcome in a state
##Function
```{r}
rankhospital <- function(state, outcome, num = "best") {
## Read outcome data
  data <- read.csv("outcome-of-care-measures.csv", colClasses = "character")
## Check that state and outcome are valid
  if(!state %in% data$State) stop("invalid state")
  outcomes = c("heart attack","heart failure","pneumonia")
  if(!outcome %in% outcomes) stop("invalid outcome")  
## Return hospital name in that state with the given rank
## 30-day death rate
  data <- data[data$State == state,]
  if(is.numeric(num) && num%%1 == 0 && nrow(data) < num) return(NA)
  if(match(outcome,outcomes)==1) {
    col_num <- 11
  }
  else if(match(outcome,outcomes)==2) {
    col_num <- 17
  }
  else col_num <- 23
  data[data[,col_num]=="Not Available",] <- NA
  data <- data[!is.na(data[,col_num]),]
  data[,col_num] <- as.numeric(data[,col_num])
  result <- data[order(data[,col_num], data[,2]),]
  if(num == "best") n <- 1
  else if (num == "worst") n <- nrow(result)
  else if (is.numeric(num) && num%%1 == 0) n <- num
  else stop("num should be integer, best or worst")
  result[n,2]
}
```
##Test function
```{r}
rankhospital("TX", "heart failure", 4)
rankhospital("MD", "heart attack", "worst")
```
##Test incorrect input

```{r}
rankhospital("MN", "heart attack", 5000)
rankhospital("MN", "heart attack", 31.2)
```
# Ranking hospitals in all states
## Function
```{r}
rankall <- function(outcome, num = "best") {
## Read outcome data
  data <- read.csv("outcome-of-care-measures.csv", colClasses = "character")
  outcomes = c("heart attack","heart failure","pneumonia")
  if(!outcome %in% outcomes) stop("invalid outcome")
  states <- unique(data$State)
  states <- sort(states)
  result <- data.frame(hospital = character(), state = character())
  
## Check that state and outcome are valid
## For each state, find the hospital of the given rank
## Return a data frame with the hospital names and the
## (abbreviated) state name
  for (cur.state in states) {
    state.data <- data[data$State == cur.state,]
    if(is.numeric(num) && num%%1 == 0 && nrow(state.data) < num) {
      state.result <- data.frame(hospital = NA, state = cur.state)
      result <- rbind(result, state.result)
    }
    else {
      if(match(outcome,outcomes)==1) {
        col_num <- 11
      }
      else if(match(outcome,outcomes)==2) {
        col_num <- 17
      }
      else col_num <- 23
    
    state.data[state.data[,col_num]=="Not Available",] <- NA
    state.data <- state.data[!is.na(state.data[,col_num]),]
    state.data[,col_num] <- as.numeric(state.data[,col_num])
    temp.result <- state.data[order(state.data[,col_num], state.data[,2]),]
    if(num == "best") n <- 1
    else if (num == "worst") n <- nrow(temp.result)
    else if (is.numeric(num) && num%%1 == 0) n <- num
    else stop("num should be integer, best or worst")
    state.result <- data.frame(hospital = temp.result[n,2], state = cur.state)
    result <- rbind(result, state.result)
    }
    
  }
  result
}
```
##Test function
```{r}
 head(rankall("heart attack", 20), 10)
```

```{r}
tail(rankall("pneumonia", "worst"), 3)

```

```{r}
 tail(rankall("heart failure"), 10)
```
#Quiz questions
##1
```{r}
best("SC", "heart attack")
```

##2
```{r}
best("NY", "pneumonia")
```

##3
```{r}
best("AK", "pneumonia")
```
##4
```{r}
rankhospital("NC", "heart attack", "worst")
```

##5
```{r}
rankhospital("WA", "heart attack", 7)
```
##6
```{r}
rankhospital("TX", "pneumonia", 10)
```

##7
```{r}
rankhospital("NY", "heart attack", 7)
```

##8
```{r}
r <- rankall("heart attack", 4)
as.character(subset(r, state == "HI")$hospital)
```
##9
```{r}
r <- rankall("pneumonia", "worst")
as.character(subset(r, state == "NJ")$hospital)
```
##10
```{r}
r <- rankall("heart failure", 10)
as.character(subset(r, state == "NV")$hospital)
```

